---
title: "Siulate SIR ct values for machine learning model"
author: "Mike Irvine"
date: "01/06/2022"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(eval=TRUE,collapse=TRUE,warning=FALSE)
```

This vignette exhaustively steps through the code, model checks and fitting procedure to use `virosolver` to estimate incidence curves from simulated data. This code can be adapted for your own datasets by replacing the Ct data frame, here called `example_ct_data`. Note that for your own datasets, substantial calibration and checks are advised.

```{r,eval=TRUE,message=FALSE}
library(virosolver)
library(tidyverse)
library(patchwork)
library(lazymcmc)
library(foreach)
library(doParallel)
library(here)

```

Simulation parameters,

```{r}
N <- 500000
sample_days <- seq(1,140,by=1)
sample_size <- 1000
set.seed(3938)
```


Get parameters,
 
```{r}
data(example_seir_partab)
pars <- example_seir_partab$values
names(pars) <- example_seir_partab$names
pars["t0"] <- 0
```

simulate SEIR model,

```{r}


times <- seq(1,200,by=1)

## Pull parameters for SEIR model
seir_pars <- c(pars["R0"]*(1/pars["infectious"]),1/pars["incubation"],1/pars["infectious"])
## Set up initial conditions.
## Note if population_n=1, then solves everything per capita
init <- c((1-pars["I0"])*N,0,pars["I0"]*N,0,0,0)

## Solve the SEIR model using the rlsoda package
#sol <- rlsoda::rlsoda(init, times, C_SEIR_model_rlsoda, parms=seir_pars, dllname="virosolver",
#                      deSolve_compatible = TRUE,return_time=TRUE,return_initial=TRUE,atol=1e-10,rtol=1e-10)

## Solve the SEIR model using the lsoda package. lsoda runs about 4x slower than rlsoda, but
## the lsoda package is available on CRAN, making it more user-friendly.
sol <- deSolve::ode(init, times, func="SEIR_model_lsoda",parms=seir_pars,
                    dllname="virosolver",initfunc="initmodSEIR",
                    nout=0, rtol=1e-6,atol=1e-6)

## Convert to data frame and column names
sol <- as.data.frame(sol)
colnames(sol) <- c("time","S","E","I","R","cumu_exposed","cumu_infected")
## Get Rt
## 
sol$Rt <- (sol$S) * pars["R0"] 

## Shift time for start
sol$time <- sol$time + floor(pars["t0"])

## Dummy rows from pre-seeding
if(pars["t0"] > 0){
  dummy_row <- data.frame("time"=0:(floor(unname(pars["t0"]))-1),"S"=N,"E"=0,"I"=0,"R"=0,"cumu_exposed"=0,"cumu_infected"=0,"Rt"=unname(pars["R0"])*N)
  sol <- bind_rows(dummy_row, sol)
}
sol <- sol[sol$time %in% times,]

## Pull raw incidence in absolute numbers
inc <- c(0,diff(sol[,"cumu_exposed"]))
inc <- pmax(0, inc)

## Get per capita incidence and prevalence
per_cap_inc <- inc/N
per_cap_prev <- (sol$E + sol$I)/N

## Melt solution (wide to long format) and get per capita
sol <- reshape2::melt(sol, id.vars="time")
sol$value <- sol$value/N

## Plot all compartments
p <- ggplot(sol) +
  geom_line(aes(x=time,y=value,col=variable)) +
  ylab("Per capita") +
  xlab("Date") +
  theme_bw()

## Plot incidence and prevalence
p_inc <- ggplot(data.frame(x=times,y=per_cap_inc,y1=per_cap_prev)) +
  geom_line(aes(x=x,y=y),col="red") +
  geom_line(aes(x=x,y=y1),col="blue") +
  ylab("Per capita incidence (red)\n and prevalence (blue)") +
  xlab("Date") +
  theme_bw()

```

Show output of SEIR model,

```{r}
show(p)
show(p_inc)
```

Next simulate infection times,

```{r}
prob_infection <- sol %>% filter(variable == "I") %>% pull(value)
infection_times <- simulate_infection_times(N,prob_infection,overall_prob = 1)
```

We now have a vector of times when infected. We can then use this to generate
for each day a sample of times since infected. For example we sample the population
randomly on a given day $t = 50$,

```{r}
# get sample day
sample_t <- 100
# sample 100 from population:
sample_from_population <- sample(infection_times,100)
times_since_infected <- if_else(sample_t > sample_from_population, 
                                sample_t - sample_from_population,
                                -1
                                )
tibble(y=times_since_infected) %>%
  ggplot(aes(x="Sample Time",y=y)) +
    geom_violin() +
    geom_jitter()
```

We can vectorize this to get samples on different days,

```{r}

sampled_times_since_infected <- tibble()
for(sample_t in sample_days){
  # sample 100 from population:
  sample_from_population <- sample(infection_times,sample_size)
  times_since_infected <- if_else(sample_t > sample_from_population, 
                                  sample_t - sample_from_population,
                                  -1
                                  )
  sampled_times_since_infected <- sampled_times_since_infected %>% 
    bind_rows(
      tibble(times_since_infected = times_since_infected,
             sample_t = sample_t, Rt = sol %>% filter(variable == "Rt",time == {{sample_t}}) %>% pull(value))
    )
}
```

Check this distribution of times since infected at each sample time,

```{r, eval=TRUE, fig.width=7,height=2.5,dpi=100}
## Plot only detectable Ct values
p_infected_times_data <- sampled_times_since_infected %>%
  filter(sample_t %in% seq(20,160,by=10)) %>% 
  ggplot() + 
  geom_violin(aes(x=sample_t,group=sample_t,y=times_since_infected),
              scale="width",fill="grey70",
              draw_quantiles=c(0.025,0.5,0.975)) + 
  geom_jitter(aes(x=sample_t,y=times_since_infected),size=0.1,width=2,height=0) + 
  #scale_y_continuous(trans="reverse") +
  theme_bw() +
  ylab("Times since infected") +
  xlab("Observation time") +
  ggtitle("")

show(p_infected_times_data)

```

From these times since infected simulate viral loads (Ct),

```{r}
# viral_loads <- simulate_viral_loads_example(sampled_times_since_infected$times_since_infected,
#                                             pars,N=1)

times_since_infected <- sampled_times_since_infected$times_since_infected

t_switch <-  pars["t_switch"] + pars["desired_mode"] + pars["tshift"]
sd_mod <- rep(pars["sd_mod"], max(times_since_infected))
unmod_vec <- 1:min(t_switch,max(times_since_infected))
sd_mod[unmod_vec] <- 1
decrease_vec <- (t_switch+1):(t_switch+pars["sd_mod_wane"])
## For the next sd_mod_wane days, variance about Ct trajectory decreases linearly
sd_mod[decrease_vec] <- 1 - ((1-pars["sd_mod"])/pars["sd_mod_wane"])*seq_len(pars["sd_mod_wane"])

sampled_ct <- matrix(ncol=1,nrow=length(times_since_infected))
for(i in 1:length(times_since_infected)){
  age <- times_since_infected[i]
  ## For each value we're going to simulate, pre-determine if it will still be detectable or not
  detectable_statuses <- rnbinom(1, 1, prob=pars["prob_detect"]) + 
    pars["tshift"] + pars["desired_mode"] + pars["t_switch"]
  cts <- rep(virosolver::viral_load_func(pars, age, FALSE, 0),1)
  cts[detectable_statuses <= age] <- 1000
  sd_used <- pars["obs_sd"]*sd_mod[age]
  ## Generate a observations of Ct values from gumbel distribution for a specified mode
  ct_obs_sim <- extraDistr::rgumbel(1, cts, sd_used)
  ## Set Ct values greater than intercept to intercept value
  ct_obs_sim <- pmin(ct_obs_sim, pars["intercept"])
  sampled_ct[i,] <- ct_obs_sim
}
sampled_ct <- data.frame(sampled_ct)

sampled_ct$time_since_infection <- times_since_infected



sampled_ct <- sampled_times_since_infected %>%
  bind_cols(sampled_ct %>% select(-time_since_infection)) 
  
```

Plot resulting Ct values,

```{r, eval=TRUE, fig.width=7,height=2.5,dpi=100}
## Plot only detectable Ct values
p_infected_times_data <- sampled_ct %>% 
  filter(sampled_ct < 40) %>% 
  filter(sample_t %in% seq(20,160,by=10)) %>% 
  ggplot() + 
  geom_violin(aes(x=sample_t,group=sample_t,y=sampled_ct),
              scale="width",fill="grey70",
              draw_quantiles=c(0.025,0.5,0.975)) + 
  geom_jitter(aes(x=sample_t,y=sampled_ct),size=0.1,width=2,height=0) + 
  scale_y_continuous(trans="reverse") +
  theme_bw() +
  ylab("Ct value") +
  xlab("Observation time") +
  ggtitle("Observed Ct values < 40 (the limit of detection) over time")

show(p_infected_times_data)

```

Generate csv file output,

```{r}
sampled_ct %>% 
  rename("time" = "sample_t","ct" = "sampled_ct") %>%
  select(time,ct,Rt) %>% 
  write_csv(here("simulated_ct_output_test.csv"))
```
